\thesistitle

### 2019-08-24 main.tex
Experimental setup
Stuff
T3R. CloneRefactor + Image
Various improvements

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}
\usepackage{booktabs}
\usepackage{paralist}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{adjustbox}
\usepackage{titlepic}
\usepackage{vhistory}
\usepackage{enumitem}
\usepackage{longtable}
\usepackage[backend=bibtex, style=ieee,citestyle=numeric-comp]{biblatex}
\usepackage[colorinlistoftodos, prependcaption]{todonotes}
\usepackage{placeins}
\usepackage{multirow}
\usepackage{amsmath,amssymb}
\usepackage[hidelinks]{hyperref}
\usepackage[acronym,nogroupskip,nohypertypes={acronym},toc,section=chapter]{glossaries}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{cmap}
\usepackage{float}
\usepackage{hyperref}
\usepackage{wrapfig}
\usepackage{titlesec}
\usepackage{listings}
\usepackage{parcolumns}
\usepackage[T1]{fontenc}
\usepackage{inconsolata}

@@

\setcounter{secnumdepth}{4}

@@

\renewcommand{\maketitle}{%
	% The cover page
	% --------------
	\thispagestyle{empty}
	\enlargethispage{30pt}
	\renewcommand{\thefootnote}{\fnsymbol{footnote}}
	% Will be page 0, s.t. contents start on page 1
	\setcounter{page}{0}
	% \eccoverhead
	% Volume and article title, author(s)
	\vspace{60pt}
	\begin{center}
		{\Huge\bfseries Towards Automatic Improvement of Software Maintainability through Automated Refactoring of Code Clones \par}
		% Earlier versions:
		% Towards Automated Refactoring of Code Clones in Object-Oriented Programming Languages
		\vspace{44pt}
		{\Large\bfseries Simon Baars \par}
		\email{simon.mailadres@gmail.com}\\
		\vspace{11pt}
		30 June 2019, \pageref{LastPage} pages
	\end{center}
	\vfill
	\begin{tabular}{ll}
		\textbf{Research supervisor:} Dr. Ana Oprescu, \email{ana.oprescu@uva.nl} \\
		\textbf{Host/Daily supervisor:} Xander Schrijen Msc., \email{x.schrijen@sig.eu}\\
		\textbf{Host organisation/Research group:} Software Improvement Group (SIG), \url{http://sig.eu/}
	\end{tabular}
	\uvacoverfoot
	\newpage
	\setcounter{footnote}{0}
	\renewcommand{\thefootnote}{\arabic{footnote}}
	\setlength{\parskip}{0pt}
}

@@

\maketitle
\input{sections/abstract}
\tableofcontents
\input{sections/introduction}
\input{sections/background}
\input{sections/research}
\input{sections/clonerefactor}
\input{sections/results}
\input{sections/discussion}
\input{sections/related_work}
\input{sections/conclusion}
\input{sections/acknowledgements}

>>>>

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}
\usepackage{booktabs}
\usepackage{paralist}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{adjustbox}
\usepackage{titlepic}
\usepackage{vhistory}
\usepackage{enumitem}
\usepackage{longtable}
\usepackage[style=ieee,citestyle=numeric-comp]{biblatex}
\usepackage[colorinlistoftodos, prependcaption]{todonotes}
\usepackage{placeins}
\usepackage{multirow}
\usepackage{amsmath,amssymb}
\usepackage[hidelinks]{hyperref}
\usepackage[acronym,nogroupskip,nohypertypes={acronym},toc,section=chapter]{glossaries}
\usepackage{amssymb}
\usepackage{verbatim}
\usepackage{cmap}
\usepackage{float}
\usepackage{hyperref}
\usepackage{wrapfig}
\usepackage{titlesec}
\usepackage{parcolumns}
\usepackage[T1]{fontenc}
\usepackage{inconsolata}
\usepackage{longtable}

@@

\setcounter{secnumdepth}{3}

@@

\renewcommand{\maketitle}{%
	% The cover page
	% --------------
	\thispagestyle{empty}
	\enlargethispage{30pt}
	\renewcommand{\thefootnote}{\fnsymbol{footnote}}
	% Will be page 0, s.t. contents start on page 1
	\setcounter{page}{0}
	% \eccoverhead
	% Volume and article title, author(s)
	\vspace{60pt}
	\begin{center}
		{\Huge\bfseries Improving Software Maintainability through Automated Refactoring of Code Clones \par}
		% Earlier versions:
		% Towards Automated Refactoring of Code Clones in Object-Oriented Programming Languages
    % Towards Automated Improvement of Software Maintainability through Automated Refactoring of Code Clones
		\vspace{44pt}
		{\Large\bfseries Simon Baars \par}
		\email{simon.mailadres@gmail.com}\\
		\vspace{11pt}
		30 June 2019, \pageref{LastPage} pages
	\end{center}
	\vfill
	\begin{tabular}{ll}
		\textbf{Research supervisor:} Dr. Ana Oprescu, \email{ana.oprescu@uva.nl} \\
		\textbf{Host/Daily supervisor:} Xander Schrijen Msc., \email{x.schrijen@sig.eu}\\
		\textbf{Host organisation/Research group:} Software Improvement Group (SIG), \url{http://sig.eu/}
	\end{tabular}
	\uvacoverfoot
	\newpage
	\setcounter{footnote}{0}
	\renewcommand{\thefootnote}{\arabic{footnote}}
	\setlength{\parskip}{0pt}
}

@@

\maketitle
\input{sections/abstract}
\tableofcontents
\input{sections/introduction}
\input{sections/background}
\input{sections/research}
\input{sections/clonerefactor}
\input{sections/experimentalsetup}
\input{sections/results}
\input{sections/discussion}
\input{sections/related_work}
\input{sections/conclusion}
\input{sections/acknowledgements}

### 2019-08-24 sections/clonerefactor.tex

Every clone class only has a single relation, which is the first relation from the above list that the clone class applies to. For instance: all ``Superclass'' clones also apply to ``Same Hierarchy'', but because ``Superclass'' is earlier in above list they will get the ``Superclass'' relation. This is because the items earlier in the list denote a more favorable refactoring.

To check a the relation for a given clone class, we compare each node in a clone instance with the respective nodes in each other clone instance within a clone class. So for a clone class with x clone instances and each clone instance has y nodes, we perform:

@@

Cloned nodes are flagged as \textit{Same method} if these nodes are found in the same method. We define the method as the first method declaration that is encountered when following the parent nodes in the ast for a given cloned node. Please note that a clone instance may not always be in a method, for which this predicate will fail.

Cloned nodes are flagged as \textit{Same class} if these nodes are found in the same class. We define the class as the first class declaration that is encountered when following the parent nodes in the ast for a given cloned node.

@@

The greatest downside on system design of putting functionality in interfaces is that interfaces are per definition part of a classes' public contract. That is, all functionality that is shared between classes via an interface cannot be hidden by setting a stricter visibility. Because of that, we favor all ``Common Hierarchy'' refactoring opportunities over ``Common Interface''.

@@

\subsubsection{Unrelated}
Clones are unrelated if they share no common class or interface in their inheritance structure. These clones are least favorable when looking at refactoring, because their refactoring will almost always have a major impact on system design. We formulated four categories of unrelated clones to look into their refactoring opportunities.

Cloned nodes are flagged \textit{No Direct Superclass} if they are in classes that do currently not have a parent. This marks the opportunity for creating a superclass abstraction and placing the extracted method there. Cloned nodes are flagged \textit{No Indirect Superclass} if any of their ancestors does not have a parent. In such a case, it would be possible to create such an abstraction for the ancestor that does not have a parent.

@@

\subsection{Location}\label{sec:setuplocation}
A paper by Lozano et al. \cite{lozano2007evaluating} discusses the harmfulness of cloning. The authors argue that 98\% are produced at method-level. However, this claim is based on a small dataset and based on human copy-paste behavior rather than static code analysis. We decided to measure the locations of clones through static analysis on our dataset. We chose the following categories:
\begin{enumerate}
  \item \textbf{Method/Constructor Level:} A clone instance that does not exceed the boundaries of a single method or constructor (optionally including the declaration of the method or constructor itself).
  \item \textbf{Class Level:} A clone instance in a class, that exceeds the boundaries of a single method or contains something else in the class (like field declarations, other methods, etc.).
  \item \textbf{Interface Level:} A clone that is (a part of) an interface.
  \item \textbf{Enumeration Level:} A clone that is (a part of) an enumeration.
\end{enumerate}
We check the location of each clone instance for each of its nodes. If any node reports a different location from the others, we choose the location that is lowest in above list. So for instance, if a clone instance has 15 nodes that denote a \textit{Method Level} location but 3 nodes are \textit{Class Level}, the clone instance becomes \textit{Class Level}.

\subsubsection{Method/Constructor Level Clones} \label{sec:methodlevelcr}
Method/Constructor Level clones denote clones that span are found in either a method or contructor. A constructor is a special method that is called when an object is instantiated. Current clone refactoring studies only focus on clones at method level \cite{choi2011extracting, yue2018automatic, kodhai2013method, arcelli2013software, lin2014clonepedia, mandal2014automatic, balazinska2000advanced, yongting2018detection, bouktif2006novel, fanqi2014using, devi2016study}. This is because most clones reside at those places \cite{lozano2007evaluating, fontana2015duplicated} and most of those clones can be refactored with a relatively simple set of refactoring techniques \cite{kodhai2013method, fontana2015duplicated}.

@@

Where \textit{T} is the type of a node. \textit{ancestors(n)} denotes all of ancestor nodes of a node. The ancestor nodes are recursively all the parents of a given node in the AST. Apart from this, we check that the MethodDeclaration ancestor of each clone instance is the same, to be sure the clone only spans a \textit{single} method. Figure \ref{fig:methodlevelclone} shows an example of a method level clone and its recursive parents.

@@

\subsubsection{Class/Interface/Enumeration Level Clones}
Class/Interface/Enumeration Level clone instances are found inside the body of one of these declarations and optionally include the declaration itself. It can also be a clone instance that exceeds the boundaries of a single method. These clone instances can contain a fields, (abstract) methods, inner classes, enumeration fields, etc. These types of clones require various refactoring techniques to refactor. For instance, we might have to move fields in a inheritance hierarchy. Or, we might have to perform a refactoring on more of an architectural level, if a large set of methods is cloned.

@@

\subsection{Contents}\label{sec:setupcontents}
Finally, we looked at what nodes individual clone instances span. We selected a set of catagories on the basis of empirical evaluation of a set of clones in our dataset. We selected the following categories to be relevant for refactoring:
\begin{enumerate}
  \item \textbf{Full Method/Constructor/Class/Interface/Enumeration:} A clone that spans a full class, method, constructor, interface or enumeration, including its declaration.
  \item \textbf{Partial Method/Constructor:} A clone that spans (a part of) the body of a method/constructor. The declaration itself is not included.
  \item \textbf{Several Methods:} A clone that spans over two or more methods, either fully or partially, but does not span anything but methods (so not fields or anything in between).
  \item \textbf{Only Fields:} A clone that spans only global variables.
  \item \textbf{Other:} Anything that does not match with above-stated categories.
\end{enumerate}

@@

\subsubsection{Partial Method/Constructor}
These categories describe clone instances which are found in the body of a method or constructor. These clones can often be refactored by extracting a new method out of the cloned code. We detect such clones by checking that each node in the clone instance has a ancestor node that is a MethodDeclaration, but none of the nodes is a method declaration. We describe the procedure by with we recursively seek the parent nodes of a node in Section~\ref{sec:methodlevelcr}.

@@

\subsubsection{Other}
The ``Other'' category denotes all configurations of clone contents that do not fall into above categories. Often, these are combinations of above states concepts. For instance, a combination of constructors and methods is clones, or a combination of fields and methods. Such clones indicate, like the ``Several Methods'', the requirement of performing a more architectural-level refactoring. These are often more complicated, especially when aiming to automate this process.

@@

\subsection{Method extraction opportunities}\label{sec:refactorabilitysetup}
The most used technique to refactor clones is ``Extract Method'' (creating a new method on the basis of the contents of clones). However, method extraction cannot be applied in all cases. In some instances, more conditions may apply to be able to conduct a refactoring, if beneficial at all. CloneRefactor maps to what extend this refactoring method can be used to refactor the found clones automatically.

@@

\subsubsection{Is Not A Partial Method}
The ``Extract Method'' refactoring can only refactor code in the body of a method. We mapped the contents of clones in Section~\ref{sec:setuplocation}. Here we defined the ``Partial Method'' catagory as \textit{a clone that spans (a part of) the body of a method. The declaration itself is not included.} If the clone does not fall within this category, ``Extract Method'' is likely not to be correct technique for refactoring such clones.

\subsubsection{Top-level Node Is Not A Statement}
A method body can only consist of statements. We cannot put declarations in a method. However, there are possibilities where a clone consists of something else than a statement as top level node. The most prominent example is when a clause is cloned (see Section~\ref{sec:astbackground} for examples of such concepts). For instance, we cannot extract only the catch-clause of a try-statement to a new method. In such a case, we should either extract the complete try-statement, or only the contents of the catch clause.

\subsubsection{Overlaps}
We do not consider clone classes for refactoring that overlap in the clone class itself. That is because such clones may require other techniques to refactor them. We consider analyzing such refactoring methods future work. An example refactoring of clone instances that overlap within their clone class is displayed in Figure~\ref{fig:overlaprefactoring}. In this example one clone instance is found between line \ref{line:overlaprefactoring1} and \ref{line:overlaprefactoring4} and the other between \ref{line:overlaprefactoring2} and \ref{line:overlaprefactoring5}. To refactor such clones, other techniques may be required, like introducing a new for-loop to reduce the duplication.

@@

\subsubsection{Can be extracted}
All clones that do not adhere to any of above-stated categories, are considered suitable for automated method extraction.

\subsection{Applying refactorings} \label{sec:applyingrefactorings}
All clones that are marked as ``Can be Extracted'' will then be refactored. The refactoring techniques that we use depend on the relation and contents of the clone. In this section, we explain the transformations we apply to the program AST in order to refactor clones.

First, we create a new method and give it an uniquely generated name. We replace all nodes from the clone instances of the clone to method calls of the newly created method. We place a copy the nodes that were removed from a single clone instance in the new method.

@@

\paragraph{Unrelated}
If all clone instances in a clone class are in different classes that do not yet have a superclass, or their ancestors have no superclass, we create a superclass abstraction for them. This way we make the implicit dependency between both clone fragments explicit, while provisioning a common place for common functionality. We give the method a \texttt{protected} visibility, which indicates that the method can only be used within its hierarchy.

@@

For unrelated clones, we always create a new abstraction instead of introducing a ``utility class''. Most current code clone refactoring support tools use utility classes to refactor unrelated code clones \cite{mazinanian2016jdeodorant, yoshida2005refactoring, gode2010clone}. For this study we refrain from doing this, because the introduction of many static methods can make a codebase hard to comprehend \cite{xing2003pseudo}. Additionally, we can use the created abstractions for further refactorings.

@@

\paragraph{Method parameters}
We determine which variables and methods that are used in the method body are not accessible anymore at the location where the extracted method is placed. For each of these variables and called methods we add a parameter to the extracted method. We then pass the appropriate data to each method call.

\paragraph{Return value}\label{sec:refreturn}
The exracted method has a return value if one of these three conditions is met:
\begin{itemize}
\item \textbf{There is a return statement in the extracted fragment}: If the extracted code fragment ends in a return statement, we set the type of the returned variable as the type of the extracted method. We turn all method calls of the extracted method into return statements followed by the method call.
\item \textbf{A method parameter is re-assigned in the extracted fragment}: If a method parameter is re-assigned, we must return its changed value. We add a return statement to the end of the body of the extracted method returning the variable that was changed. We set the return type of the extracted method to the type of the variable that was assigned. We turn all method calls of the extracted method into variable assignment statements in which we assign the changed variable.
\item \textbf{A variable is declared in the extracted fragment}: If a variable is declared in the extracted fragment, but used outside it, we return the declared variable. We add a return statement to the end of the body of the extracted method that returns the declared variable. If the variable declaration statement is the final statement in the method body, we remove the variable declaration and directly return the value it assigns to the declared variable. We set the return type of the extracted method to the type of the variable that was declared. We turn all method calls of the extracted method into variable declaration statements in which we declare the variable returned by the extracted method.
\end{itemize}

\paragraph{Thrown exception}
If a method in Java throws an exception that is not a \texttt{RuntimeException}, it must be denoted in the method declaration. Because of this, we analyze the extracted method for \texttt{throw} statements, which denote an exception being thrown. We also analyze all called methods to find out whether they throw exceptions. For each of these exception, we check whether they are caught in the extracted method. Any exception that is that is not a \texttt{RuntimeException} and that is not being caught is added to the thrown exceptions of the extracted method.

@@

In this case we are dealing with single refactorings and we want to measure the maintainability impact of such a small change. Because of that, we measure only a subset of the metrics \cite{heitlager2007practical} and focus on the absolute metric changes (instead of the risk profiles). The subset of metrics we decided to focus on are all metrics that are measured on method level (as the other metrics show a lesser impact on the maintainability for these small changes). These metrics are
\begin{itemize}
\item \textbf{Duplication}: In Heitlager et al. \cite{heitlager2007practical} this metric is measured by taking the amount of duplicated lines. We decided to use the amount of duplicated tokens part of a clone class instead, to have a stronger reflection of the impact of the refactoring by measuring a more fine-grained system property.
\item \textbf{Volume}: The more code a system has, the more code has to be maintained. The paper \cite{heitlager2007practical} measures volume as lines of code. As with duplication, we use the amount of tokens instead.
\item \textbf{Complexity}: Heitlager et al. use MCCabe complexity \cite{mccabe1976complexity} to calculate their complexity metric. The MCCabe complexity is a quantative measure of the number of linearly independent paths through a method.
\item \textbf{Method Interface Size}: The amount of parameters that a method has. If a method has many parameters, the code may become harder to understand and it is an indication that data is not encapsulated adequately in objects \cite{fowler2018refactoring}.
\end{itemize}

@@

\begin{itemize}
\item \textbf{Duplication}: The duplication can only decrease as a result of the refactoring. We calculate the reduction of duplication as the product of the number of tokens in the body of the extracted and the number of clone instances in the clone class.
\item \textbf{Volume}: The volume can both decrease and increase as a result of the refactoring. The reduction of volume is equal to the reduction of duplication. The volume is increased by the amount of tokens in each method call to the extracted method, plus optionally the return, assignment or declaration of the extracted method. We also add the size in tokens of the extracted method itself to the volume. If the extracted method is moved to a newly created class or interface, we use volume of that declaration instead (because it in turn contains the extracted method). The added volume minus the reduced volume results in the change in volume.
\item \textbf{Complexity}: The complexity can both decrease and increase as a result of the refactoring. To calculate the reduction of complexity, we take the sum of all branching statements in all clone instances. The increase in complexity is the complexity of the extracted method (the number of branches plus one). The added volume complexity the reduced complexity results in the change in complexity.
\item \textbf{Method Interface Size}: The duplication can only increase as a result of the refactoring. The method interface size increases by the amount of parameters in the extracted method.
\end{itemize}

>>>>

Every clone class only has a single relation, which is the first relation from the above list that the clone class applies to. For instance: all ``Superclass'' clones also apply to ``Same Hierarchy'', but because ``Superclass'' is earlier in the list they will get the ``Superclass'' relation. This is because the items earlier in the list denote a more favorable refactoring.

To determine the relation for a given clone class, we compare each node in a clone instance with the respective nodes in each other clone instance within a clone class. So for a clone class with x clone instances and each clone instance has y nodes, we perform:

@@

Cloned nodes are flagged as \textit{Same method} if these nodes are found in the same method. We define the method as the first method declaration that is encountered when following the parent nodes in the AST for a given cloned node. Please note that a clone instance may not always be in a method, for which this predicate will fail.

Cloned nodes are flagged as \textit{Same class} if these nodes are found in the same class. We define the class as the first class declaration that is encountered when following the parent nodes in the AST for a given cloned node.

@@

The greatest downside on system design of putting functionality in interfaces is that interfaces are per definition part of a classes' public contract. That is, all functionality that is shared between classes via an interface cannot be hidden by stricter visibility. Because of that, we favor all ``Common Hierarchy'' refactoring opportunities over ``Common Interface''.

@@

\subsubsection{Unrelated}
Clones are unrelated if they share no common class or interface in their inheritance structure. These clones are least favorable for refactoring, because their refactoring will almost always have a major impact on system design. We formulated four categories of unrelated clones to look into their refactoring opportunities.

Cloned nodes are flagged \textit{No Direct Superclass} if they are in classes that do currently not have a parent. This marks the opportunity for creating a superclass abstraction and placing the extracted method there. Cloned nodes are flagged \textit{No Indirect Superclass} if any of their ancestors do not have a parent. In such a case, it would be possible to create such an abstraction for the ancestor that does not have a parent.

@@

\subsection{Location}\label{sec:setuplocation}
A paper by Lozano et al. \cite{lozano2007evaluating} discusses the harmfulness of cloning. The authors argue that 98\% are produced at method-level. However, this claim is based on a small dataset and based on human copy-paste behavior rather than static code analysis. We decided to measure the locations of clones through static analysis on our dataset. We chose the following categories:
\begin{enumerate}
  \item \textbf{Method/Constructor Level:} A clone instance that does not exceed the boundaries of a single method or constructor (optionally including the declaration of the method or constructor itself).
  \item \textbf{Class Level:} A clone instance in a class, that exceeds the boundaries of a single method or contains something else in the class (like field declarations, other methods, etc.).
  \item \textbf{Interface Level:} A clone that is (a part of) an interface.
  \item \textbf{Enumeration Level:} A clone that is (a part of) an enumeration.
\end{enumerate}
We check the location of each clone instance for each of its nodes. If any node reports a different location from the others, we choose the location that is lowest in the above list. So for instance, if a clone instance has 15 nodes that denote a \textit{Method Level} location but 3 nodes are \textit{Class Level}, the clone instance becomes \textit{Class Level}.

\subsubsection{Method/Constructor Level Clones} \label{sec:methodlevelcr}
Method/Constructor Level clones denote clones that are found in either a method or constructor. A constructor is a special method that is called when an object is instantiated. Current clone refactoring studies only focus on clones at method level \cite{choi2011extracting, yue2018automatic, kodhai2013method, arcelli2013software, lin2014clonepedia, mandal2014automatic, balazinska2000advanced, yongting2018detection, bouktif2006novel, fanqi2014using, devi2016study}. This is because most clones reside at those places \cite{lozano2007evaluating, fontana2015duplicated} and most of those clones can be refactored with a relatively simple set of refactoring techniques \cite{kodhai2013method, fontana2015duplicated}.

@@

Where \textit{T} is the type of a node. \textit{ancestors(n)} denotes all of ancestor nodes of a node. The ancestor nodes are all recursive parents of a given node in the AST. Apart from this, we check that the MethodDeclaration ancestor of each clone instance is the same, to be sure the clone only spans a \textit{single} method. Figure \ref{fig:methodlevelclone} shows an example of a method level clone and its recursive parents.

@@

\subsubsection{Class/Interface/Enumeration Level Clones}
Class/Interface/Enumeration Level clone instances are found inside the body of one of these declarations and optionally include the declaration itself. It can also be a clone instance that exceeds the boundaries of a single method. These clone instances can contain fields, (abstract) methods, inner classes, enumeration fields, etc. These types of clones require various refactoring techniques to refactor. For instance, we might have to move fields in an inheritance hierarchy. Or, we might have to perform a refactoring on more of an architectural level, if a large set of methods is cloned.

@@

\subsection{Contents}\label{sec:setupcontents}
Finally, we looked at what nodes individual clone instances span. We selected a set of categories based on empirical evaluation of a set of clones in our dataset. We selected the following categories to be relevant for refactoring:
\begin{enumerate}
  \item \textbf{Full Method/Constructor/Class/Interface/Enumeration:} A clone that spans a full class, method, constructor, interface or enumeration, including its declaration.
  \item \textbf{Partial Method/Constructor:} A clone that spans (a part of) the body of a method/constructor. The declaration itself is not included.
  \item \textbf{Several Methods:} A clone that spans over two or more methods, either fully or partially, but does not span anything but methods (so not fields or anything in between).
  \item \textbf{Only Fields:} A clone that spans only global variables.
  \item \textbf{Other:} Anything that does not match with above-stated categories.
\end{enumerate}

@@

\subsubsection{Partial Method/Constructor}
These categories describe clone instances which are found in the body of a method or constructor. These clones can often be refactored by extracting a new method out of the cloned code. We detect such clones by checking that each node in the clone instance has an ancestor node that is a MethodDeclaration, but none of the nodes is a method declaration. We describe the procedure by with we recursively seek the parent nodes of a node in Section~\ref{sec:methodlevelcr}.

@@

\subsubsection{Other}
The ``Other'' category denotes all configurations of clone contents that do not fall into above categories. Often, these are combinations of the above stated concepts. For instance, a combination of constructors and methods or a combination of fields and methods is cloned. Such clones indicate, like the ``Several Methods'', the requirement of performing a more architectural-level refactoring. These are often more complicated, especially when aiming to automate this process.

@@

\subsection{Method extraction opportunities}\label{sec:refactorabilitysetup}
The most used technique to refactor clones is ``Extract Method'' (creating a new method based on the contents of clones). However, method extraction cannot be applied in all cases. In some instances, more conditions may apply to be able to conduct a refactoring, if beneficial at all. CloneRefactor maps to what extent this refactoring method can be used to refactor the found clones automatically.

@@

\subsubsection{Is Not A Partial Method}
The ``Extract Method'' refactoring can only refactor code in the body of a method. We mapped the contents of clones in Section~\ref{sec:setuplocation}. Here we defined the ``Partial Method'' category as \textit{a clone that spans (a part of) the body of a method. The declaration itself is not included.} If the clone does not fall within this category, ``Extract Method'' is likely not to be the correct technique for refactoring such clones.

\subsubsection{Top-level Node Is Not A Statement}
A method body can only consist of statements. We cannot put declarations in a method. However, there are possibilities where a clone consists of something else than a statement as a top-level node. Upon manual inspection, we noticed that most clones falling into this category are clones that span part of a lambda or anonymous class.

\subsubsection{Overlaps}
We do not consider clone classes for refactoring that overlap in the clone class itself. That is because such clones may require other techniques to refactor them. An example refactoring of clone instances that overlap within their clone class is displayed in Figure~\ref{fig:overlaprefactoring}. In this example one clone instance is found between line \ref{line:overlaprefactoring1} and \ref{line:overlaprefactoring4} and the other between \ref{line:overlaprefactoring2} and \ref{line:overlaprefactoring5}. To refactor such clones, other techniques may be required, like introducing a new for-loop to reduce the duplication.

@@

\subsubsection{Can be extracted}
All clones that do not adhere to any of above-stated categories are considered suitable for automated method extraction.

\subsection{Applying refactorings} \label{sec:applyingrefactorings}
All clones that are marked as ``Can be Extracted'' will then be refactored. The refactoring techniques that we use depend on the relation and contents of the clone. In this section, we explain the transformations we apply to the program AST to refactor clones.

First, we create a new method and give it a uniquely generated name. We replace all nodes from the clone instances of the clone to method calls of the newly created method. We place a copy of the nodes that were removed from a single clone instance in the new method.

@@

\paragraph{Unrelated}
If all clone instances in a clone class are in different classes that do not yet have a superclass, or their ancestors have no superclass, we create a superclass abstraction for them. This way we make the implicit dependency between both clone fragments explicit while provisioning a common place for common functionality. We give the method a \texttt{protected} visibility, which indicates that the method can only be used within its hierarchy.

@@

For unrelated clones, we always create a new abstraction instead of introducing a ``utility class''. Most current code clone refactoring support tools use utility classes to refactor unrelated code clones \cite{mazinanian2016jdeodorant, yoshida2005refactoring, gode2010clone}. For this study, we refrain from doing this because the introduction of many static methods can make a codebase hard to comprehend \cite{xing2003pseudo}. Additionally, we can use the created abstractions for further refactorings.

@@

\paragraph{Method parameters}
We determine which variables and methods that are used in the method body are not accessible anymore at the location where the extracted method is placed. For each of these variables and called methods, we add a parameter to the extracted method. We then pass the appropriate data to each method call.

\paragraph{Return value}\label{sec:refreturn}
The extracted method has a return value if one of these three conditions is met:
\begin{itemize}
\item \textbf{There is a return statement in the extracted fragment}: If the extracted code fragment ends in a return statement, we change the type of the returned extracted method to the type of the returned variable. We turn all method calls of the extracted method into return statements followed by the method call.
\item \textbf{A method parameter is re-assigned in the extracted fragment}: If a method parameter is re-assigned, we must return its changed value. We add a return statement to the end of the body of the extracted method returning the variable that was changed. We set the return type of the extracted method to the type of the variable that was assigned. We turn all method calls of the extracted method into variable assignment statements in which we assign the changed variable.
\item \textbf{A variable is declared in the extracted fragment}: If a variable is declared in the extracted fragment, but used outside it, we return the declared variable. We add a return statement to the end of the body of the extracted method that returns the declared variable. If the variable declaration statement is the final statement in the method body, we remove the variable declaration and directly return the value it assigns to the declared variable. We set the return type of the extracted method to the type of the variable that was declared. We turn all method calls of the extracted method into variable declaration statements in which we declare the variable returned by the extracted method.
\end{itemize}

\paragraph{Thrown exception}
If a method in Java throws an exception that is not a \texttt{RuntimeException}, it must be denoted in the method declaration. Because of this, we analyze the extracted method for \texttt{throw} statements, which denote an exception being thrown. We also analyze all called methods to find out whether they throw exceptions. For each of these exceptions, we check whether they are caught in the extracted method. Any exception that is not a \texttt{RuntimeException} and that is not being caught is added to the thrown exceptions of the extracted method.

@@

In this case, we are dealing with single refactorings and we want to measure the maintainability impact of such a small change. Because of that, we measure only a subset of the metrics \cite{heitlager2007practical} and focus on the absolute metric changes (instead of the risk profiles). The subset of metrics we decided to focus on are all metrics that are measured on method level (as the other metrics show a lesser impact on the maintainability for these small changes). These metrics are
\begin{itemize}
\item \textbf{Duplication}: In Heitlager et al. \cite{heitlager2007practical} this metric is measured by taking the amount of duplicated lines. We decided to use the amount of duplicated tokens part of a clone class instead, to have a stronger reflection of the impact of the refactoring by measuring a more fine-grained system property.
\item \textbf{Volume}: The more code a system has, the more code has to be maintained. The paper \cite{heitlager2007practical} measures volume as lines of code. As with duplication, we use the number of tokens instead.
\item \textbf{Complexity}: Heitlager et al. use MCCabe complexity \cite{mccabe1976complexity} to calculate their complexity metric. The MCCabe complexity is a quantitative measure of the number of linearly independent paths through a method.
\item \textbf{Method Interface Size}: The number of parameters that a method has. If a method has many parameters, the code may become harder to understand and it is an indication that data is not encapsulated adequately in objects \cite{fowler2018refactoring}.
\end{itemize}

@@

\begin{itemize}
\item \textbf{Duplication}: The duplication can only decrease as a result of the refactoring. We calculate the reduction of duplication as the product of the number of tokens in the body of the extracted and the number of clone instances in the clone class.
\item \textbf{Volume}: The volume can both decrease and increase as a result of the refactoring. The reduction of the volume is equal to the reduction of duplication. The volume is increased by the number of tokens in each method call to the extracted method, plus optionally the return, assignment or declaration of the extracted method. We also add the size in tokens of the extracted method itself to the volume. If the extracted method is moved to a newly created class or interface, we use the volume of that declaration instead (because it contains the extracted method). The added volume minus the reduced volume results in the change in volume.
\item \textbf{Complexity}: The complexity can both decrease and increase as a result of the refactoring. To calculate the reduction of complexity, we take the sum of all branching statements in all clone instances. The increase in complexity is the complexity of the extracted method (the number of branches plus one). The added volume complexity the reduced complexity results in the change in complexity.
\item \textbf{Method Interface Size}: The duplication can only increase as a result of the refactoring. The method interface size increases by the number of parameters in the extracted method.
\end{itemize}

### 2019-08-24 sections/discussion.tex

\subsubsection{Refactorability} \label{sec:discussrefactorability}
In Section~\ref{sec:refactorability} we introduced catagories to determine the refactorability of clones through method extraction. We excluded categories that could not not be directly refactored through method extraction. However, with a few transformations or further considerations it might be possible to make these clones refactorable. In this section we will highlight a few of these categories which we believe to be refactorable through method extraction with more effort.

>>>>

\subsubsection{Refactorability} \label{sec:discussrefactorability}
In Section~\ref{sec:refactorabilitysetup} we introduced catagories to determine the refactorability of clones through method extraction. We excluded categories that could not not be directly refactored through method extraction. However, with a few transformations or further considerations it might be possible to make these clones refactorable. In this section we will highlight a few of these categories which we believe to be refactorable through method extraction with more effort.
